---
title: "Volcano Eruptions"
author: "Francis Lin | #TidyTuesday |"
date: "2020-05-12"
output: rmarkdown::github_document
---

# Introduction

Today, I played around with the `rnaturalearth` package by plotting out the locations of volcanos! The data was supplied by [The Smithsonian Institute](https://volcano.si.edu/).

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(rnaturalearth)
library(lubridate)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
volcano <- read_csv(paste0(data_path, "volcano.csv"))
```


## Manipulate data

Here I calculated the number of years elapsed since the last eruption of each volcano.

```{r manipulate data}
volcanos <- volcano %>%
    select(last_eruption_year, longitude, latitude) %>%
    mutate(years_since_eruption = year(Sys.Date())-as.numeric(last_eruption_year))
```


## Plot Data

```{r plot data}
# get country polygons
world <- ne_countries(scale = "medium", returnclass = "sf")

# plot
p <- ggplot(volcanos) +
    
    # world polygons
    geom_sf(data=world, fill = "black", colour = "grey10") +
    
    # volcano points
    geom_point(aes(x=longitude, y=latitude, color=years_since_eruption), size=1, alpha=0.9) + 
    
    # color volcano points
    scale_colour_gradient(low="red", high="yellow", na.value="white") +
    
    # aesthetics
    theme(
        panel.background = element_rect(fill = "darkslategray3"),
        plot.background = element_rect(fill = "darkslategray3"),
        legend.background = element_rect(fill = "darkslategray3"),
        panel.grid.major = element_blank(),
        plot.subtitle=element_text(size=8)
    ) +
    
    # labels
    labs(title="Years from Last Volcano Eruption", color="Years", subtitle="Points in white are unknown", x="", y="")
    
p
```

This was my first time using the `rnaturalearth` package. Although I thought it would be much more intimidating, the task ended up being easier than I thought. That's an extra tool in my toolkit now: graphing geographical locations on the world map.


## Save Image

```{r save image}
ggsave("plot/plot_2020-05-12.png", p, width=8, height=4, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
