---
title: "African American Achievements"
author: "Francis Lin | #TidyTuesday |"
date: "2020-06-09"
output: rmarkdown::github_document
---

# Introduction

[George Floyd's death](https://www.nytimes.com/2020/05/31/us/george-floyd-investigation.html) has sparked massive protests across the United States in support of #BlackLivesMatter. This week's data comes from [Wikipedia](https://en.wikipedia.org/wiki/List_of_African-American_firsts) and celebrates African American achievements. I wanted to take a look at if the percentage of female achievements increases over time.

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(scales)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
firsts <- read_csv(paste0(data_path, "firsts.csv"))
```


## Manipulate data

```{r manipulate data}
# calculate median year
median_year <- median(firsts$year)

firsts_grouped <- firsts %>%
    # factorize gender
    mutate(gender=factor(gender, levels=c("Female African American Firsts", "African-American Firsts"))) %>%
    
    # create factors based on each category-year combination
    mutate(year=ifelse(year>=median_year, paste("after", median_year), paste("before", median_year))) %>%
    unite("category", category, year, sep="\n", remove=FALSE) %>%
    mutate(category = factor(category, levels=c(sort(unique(.$category)[1:8]), sort(unique(.$category)[9:16])))) %>%
    
    # count by gender for each category
    group_by(category, year, gender) %>%
    summarize(count=n()) %>%
    
    # include calculations for pie chart
    mutate(fraction=count/sum(count),
           label=ifelse(grepl("Female", gender), paste0(round(fraction*100,1), "%"), ifelse(fraction==1, "0.0%", "")),
           ymax=cumsum(fraction),
           ymin=c(0, head(ymax, n=-1)))
```


## Plot Data

```{r plot data, fig.width=12, fig.height=5}
p <- ggplot(firsts_grouped, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=gender)) +
    
    # pie chart
    geom_rect() +
    coord_polar("y", start=0) +
    xlim(c(1, 5)) +
    
    # facet wrap per category
    facet_wrap(~ category, nrow = 2) +
    
    # labels
    labs(title="African American Achievements", fill="", subtitle="The percentage of female African American achievements have increased in all categories except Social & Jobs and Sports.") +
    geom_text(aes(label = label, x = 1, y = 0), size=3) +
    
    # aesthetics
    theme_void() + 
    theme(legend.position = "bottom",
          plot.title = element_text(margin = margin(b=5), size = 20, hjust = 0.5),
          plot.subtitle = element_text(margin = margin(b=10), size = 10, hjust = 0.5)
          )
p
```

From the image above, we can see that the percentage of female African American achievements have increased in all categories except Social & Jobs and Sports.

## Save Image

```{r save image}
ggsave("plot/plot_2020-06-09.png", p, width=12, height=5, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
