---
title: "Spotify Songs"
author: "Francis Lin | #TidyTuesday |"
date: "2020-01-21"
output: rmarkdown::github_document
---

# Introduction

The data this week comes from Spotify and the [spotifyr package](https://www.rcharlie.com/spotifyr/). In this analysis, I was curious about whether certain song attributes are correlated with the popularity of the song and how that correlation changed over time.

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(lubridate)
library(readr)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```

## Load Data

```{r load data, message=FALSE}
# read in data
spotify_songs <- read_csv(paste0(data_path, "spotify_songs.csv"))
```

## Manipulate Data

```{r manipulate data}
# find unique songs, add year
spotify_songs <- spotify_songs %>%
    distinct(track_name, track_artist, .keep_all = TRUE) %>%
    mutate(year=year(ymd(track_album_release_date, truncated=2)))

# find correlation of popularity with song attributes
spotify_corr <- spotify_songs %>%
    distinct(track_name, track_artist, .keep_all = TRUE) %>%
    mutate(year=year(ymd(track_album_release_date, truncated=2))) %>%
    group_by(year) %>%
    filter(n() >= 10) %>%
    summarise("Danceability"=cor(track_popularity, danceability),
              "Energy"=cor(track_popularity, energy),
              "Loudness"=cor(track_popularity, loudness),
              "Speechiness"=cor(track_popularity, speechiness),
              "Acousticness"=cor(track_popularity, acousticness),
              "Instrumentalness"=cor(track_popularity, instrumentalness),
              "Liveness"=cor(track_popularity, liveness),
              "Happiness (valence)"=cor(track_popularity, valence),
              "Tempo"=cor(track_popularity, tempo),
              "Song Length"=cor(track_popularity, duration_ms)) %>%
    pivot_longer(-year, names_to="Parameter", values_to="score")
```

## Plot Data

```{r plot data, message=FALSE}
p <- ggplot(spotify_corr) + 
    geom_point(aes(x=year, y=score), size=0.5, col="tomato") +
    geom_smooth(aes(x=year, y=score), se=FALSE) +
    geom_hline(yintercept=0, alpha=0.5) +
    facet_wrap(~Parameter, ncol = 5) + 
    labs(title="Correlation between Song Attributes and Popularity over Time (1965-2020)", x="Year", y="Correlation Coefficient")
p
```

From this graph, it seems we can draw a few quick conclusions:

- The tempo-popularity and liveness-popularity correlation does not change much throughout the years.
- Songs with acousticness are slowly becoming more popular as time goes on.
- Danceability of a track was positively correlated with its popularity only before 1980 and after 2010.
- The energy of a song was only negatively correlated with popularity in recent years.
- The popularity of happy songs are now on the rise.
- Instrumentalness of a song used to be positively correlated with popularity until about 1980.
- Loud songs are always popular, although less so recently.
- Longer songs tend to be slightly less popular.
- Speechiness in songs have recently taken off. This is probably due to rise in popularity of rap songs, which tend to have high speechiness ratings.

## Save Image

```{r save image, message=FALSE}
ggsave("plot/plot_2020-01-21.png", p, width=12, height=6, units="in")
```

## Session Info

```{r session info}
#session info
sessionInfo()
```
