---
title: "Broadway Weekly Grosses"
author: "Francis Lin | #TidyTuesday |"
date: "2020-04-28"
output: rmarkdown::github_document
---

# Introduction

I wanted to try my hand at some animated graphs again. This week's data comes from [Playbill](https://www.playbill.com/grosses). In this analysis, I made a racing bar chart using Broadway weekly grosses.

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(lubridate)
library(gganimate)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
grosses <- read_csv(paste0(data_path, "grosses.csv"))
```


## Manipulate data

To make a racing bar chart, I needed to interpolate between given values.

```{r manipulate data}
grosses_df <- grosses %>%
  
  # select relevant columns
  select(week_ending, show, weekly_gross) %>%
    
  # find sum of gross per week per show
  group_by(week_ending, show) %>%
  summarise(weekly_gross=sum(weekly_gross)) %>%
  ungroup() %>%
    
  # sort by week
  mutate(week_ending=ymd(week_ending)) %>%
  arrange(week_ending) %>%
    
  # calculate cumulative sum
  mutate(cumsum=ave(weekly_gross, show, FUN=cumsum))
  
grosses_df_interp <- grosses_df %>%
  # interpolate between shows and find cumulative sum
  group_by(show) %>%
  filter(n()>1) %>%
  complete(week_ending = full_seq(grosses_df$week_ending, 1)) %>%
  mutate(cumsum_smooth = (approx(x = week_ending, y = cumsum, xout = week_ending)$y)/1000000) %>%
  fill(cumsum_smooth) %>%
  
  # rank by week
  select(week_ending, show, cumsum_smooth) %>%
  group_by(week_ending) %>% 
  mutate(rank = min_rank(-cumsum_smooth) * 1) %>%
  
  # choose top 10 per week
  filter(rank <= 10) %>%
  ungroup()
```


## Plot Data

The .gif takes a long time to render.. be patient!

```{r plot data, fig.width=7, fig.height=5}
p <- ggplot(grosses_df_interp, aes(x=rank, y=show, label=show, group=show, fill = show)) +
  
  # tile
  geom_tile(aes(y=cumsum_smooth/2, height=cumsum_smooth, width=0.9, fill=show), alpha=0.8) +
  
  # label of show
  geom_text(aes(y = cumsum_smooth, label = show, hjust = ifelse(cumsum_smooth > 800, 1, 0))) +
    
  # flip coordinate
  coord_flip(clip = "off", expand = FALSE) +
    
  # display dollar amounts with commas
  scale_y_continuous(labels = scales::comma) +
    
  # display shows in descending order
  scale_x_reverse() +
    
  # aesthetics
  theme_minimal() +
  theme(plot.title=element_text(face="bold", hjust=0, size=30), axis.ticks.y=element_blank(), axis.text.y=element_blank(), panel.grid.major = element_blank(), legend.position="none") +
  labs(title='{format(as.Date(closest_state), "%Y %b %d")}', x="", y="Total Gross (Millions of Dollars)") +
  
  # animation
  transition_states(week_ending, transition_length=5, state_length=1, wrap = FALSE) +
  ease_aes("cubic-in-out")

animation <- animate(p, nframes=27000, fps=50, end_pause=100, width=546, height=390)
animation
```


## Save Image

```{r save image}
anim_save("plot/plot_2020-04-28.gif", animation)
```


## Session Info

```{r session info}
sessionInfo()
```
