---
title: "Marble Racing"
author: "Francis Lin | #TidyTuesday |"
date: "2020-06-02"
output: rmarkdown::github_document
---

# Introduction

Did you know marble racing was a thing? I had no idea. This week's data comes from [Jelle's Marble Runs](https://www.youtube.com/channel/UCYJdpnjuSWVOLgGT9fIzL0g) and contains track completion time for each marble. I wouldn't have guessed, but some marbles do consistently perform better than others. I took the opportunity to see if I can replicate [this visualization by Randal Olson](http://www.randalolson.com/2020/05/24/a-data-driven-look-at-marble-racing/).

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(scales)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
marbles <- read_csv(paste0(data_path, "marbles.csv"))
```


## Manipulate data

```{r manipulate data}
# find average race time of each race
marbles_avg_time <- marbles %>%
    group_by(race) %>%
    summarize(avg_time_race = mean(time_s, na.rm=TRUE))

# calculate time difference between each marble's time per race and the average race time
marbles_diff <- marbles %>%
    left_join(marbles_avg_time, by="race") %>%
    filter(!is.na(time_s)) %>%
    mutate(time_diff = 1+(time_s - avg_time_race)/avg_time_race)
```


## Plot Data

```{r plot data, fig.width=9, fig.height=12}
# create color palette
colorpalette = colorRampPalette(c("#5d669e", "#8dc4a6", "#f6f3c1", "#e39065", "#9a254c"))

p <- ggplot(marbles_diff, aes(x=fct_reorder(marble_name, time_diff, .fun=median, .desc=TRUE), 
                         y=time_diff
                         )) +
    stat_boxplot(geom ='errorbar', width = 0.5) + 
    geom_boxplot(fill=colorpalette(32), outlier.shape=18) +
    scale_y_continuous(labels=scales::number_format(accuracy=0.01, suffix="x")) +
    coord_flip() +
    labs(x="Marble Names", 
         y="Individual Race Times Relative to Average Race Times",
         title="Not all marble racers are created equal") +
    theme(panel.grid.minor=element_blank(),
          panel.grid.major.x=element_line(colour="grey90", size=0.5),
          panel.grid.major.y=element_blank(),
          panel.background=element_blank(),
          panel.border=element_rect(colour = "grey80", fill=NA, size=0.5),
          axis.title=element_text(size=16),
          axis.ticks=element_blank(),
          axis.text=element_text(size=12),
          plot.title = element_text(size=18, hjust = 0.5))
p
```


## Save Image

```{r save image}
ggsave("plot/plot_2020-06-02.png", p, width=9, height=12, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
