---
title: "Caribou Location Tracking"
author: "Francis Lin | #TidyTuesday |"
date: "2020-06-23"
output: rmarkdown::github_document
---

# Introduction

This week's data focuses on caribou location tracking. The data contains longitude and latitude coordinates of 260 caribou in northern British Columbia from 1988 to 2016. I didn't have too much time this week, so I created a visualization similar to the one I made for [San Francisco Trees](https://github.com/francislin96/tidytuesday/tree/master/projects/2020-01-28).

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(ggmap)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
locations <- read_csv(paste0(data_path, "locations.csv"))
```


## Create Map

```{r create map, message=FALSE}
# read in google maps API key
register_google(key=read_key("google"))

# create ggmap
caribou_map <- get_map(location = c(mean(caribou_locations$longitude), mean(caribou_locations$latitude)), zoom=7)
caribou_ggmap <- ggmap(caribou_map)
```


## Manipulate data

```{r manipulate data}
caribou_locations <- locations %>%
    
    # drop NA
    drop_na(longitude, latitude) %>%
    
    # filter for points within boundaries of map
    filter(between(longitude, min(caribou_ggmap$data$lon), max(caribou_ggmap$data$lon)), between(latitude, min(caribou_ggmap$data$lat), max(caribou_ggmap$data$lat)))
```


## Plot Data

```{r plot data, fig.width=7, fig.height=5}
p <- caribou_ggmap +
    stat_density2d(data=caribou_locations, aes(x=longitude, y=latitude, fill=season), bins=5, alpha=0.5, geom="polygon") +
    labs(x="", y="", title="Location Density of Caribou in Northern British Columbia 1988-2016", fill="Season") +
    theme(plot.title=element_text(hjust = 0.5))
p
```


## Save Image

```{r save image}
ggsave("plot/plot_2020-06-23.png", p, width=7, height=5, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
