---
title: "Australian Pets"
author: "Francis Lin | #TidyTuesday |"
date: "2020-07-14"
output: rmarkdown::github_document
---

# Introduction

It's sad seeing so many cats and dogs being euthanized in this dataset. Cats have a higher euthanization rate compared to dogs, although that gap is narrowing as both cats and dogs euthanization rates have lowered in recent years. According to [American Humane](https://americanhumane.org/fact-sheet/animal-shelter-euthanasia-2/), cats are euthanized more frequently since they are less likely to have owner identification when entering a shelter.

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(magick)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
animal_outcomes <- read_csv(paste0(data_path, "animal_outcomes.csv"))
```


## Manipulate data

```{r manipulate data}
outcomes <- animal_outcomes %>% 
    select(year, animal_type, outcome, Total) %>%
    filter(animal_type %in% c("Cats", "Dogs")) %>%
    mutate(outcome_type = factor(case_when(outcome %in% c("Reclaimed", "Rehomed") ~ "Positive",
                     outcome == "Euthanized" ~ "Negative",
                     TRUE ~ "Other"), levels=c("Positive", "Negative", "Other"))) %>%
    group_by(year, animal_type, outcome_type) %>%
    summarize(total = sum(Total)) %>%
    group_by(year, animal_type) %>%
    mutate(percentage = total/sum(total)*100) %>%
    filter(outcome_type=="Negative")
```


## Plot Data

```{r plot data, fig.width=7, fig.height=5}
# used for plotting images
imagexcoord <- 2000
dogycoord <- (outcomes %>% filter(year==imagexcoord, animal_type=="Dogs"))%>% `[[`("percentage")
catycoord <- (outcomes %>% filter(year==imagexcoord, animal_type=="Cats"))%>% `[[`("percentage")

# plot
p <- ggplot(outcomes) +
    geom_line(aes(x=year, y=percentage, color=animal_type)) +
    ylim(0,100) +
    labs(x="Year", y="Percentage Euthanized", color="Animal", title="Euthanization Rates of Cats and Dogs in Australia", subtitle="Cats have consistently been euthanized at a higher rate than dogs, although both are trending lower") + 
    theme(plot.subtitle=element_text(size=8), legend.position = "none") +
    
    # add images of dog and cat
    annotation_raster(image_read("https://cdn.pixabay.com/photo/2016/02/07/19/47/dog-1185460_960_720.png"), imagexcoord-1, imagexcoord+1, dogycoord, dogycoord+10) +
    annotation_raster(image_read("https://freesvg.org/img/molumen-cat-icons-4.png"), imagexcoord-0.9, imagexcoord+0.9, catycoord-1, catycoord+13)
p
```


## Save Image

```{r save image}
ggsave("plot/plot_2020-07-14.png", p, width=7, height=5, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
