---
title: "Philadelphia Parking Violations"
author: "Francis Lin | #TidyTuesday |"
date: "2019-12-03"
output: rmarkdown::github_document
---

# Introduction

Welcome to my first [#TidyTuesday](https://github.com/rfordatascience/tidytuesday)! This week, I'll be looking at a sample of parking violations in Philadelphia. The full dataset is available at [Open Data Philly](https://www.opendataphilly.org/dataset/parking-violations).

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(lubridate)

# set data path
data_path <- set_data_path()
```

## Load Data

```{r load data}
# read in data
tickets <- read.csv(paste(data_path, "tickets.csv", sep=""))
```

## Manipulate Data

I wanted to find the number of tickets issued by weekday. Using the pipe operator, I was able to parse the date column to obtain the weekday. I then grouped the data by weekday and counted the number of observations per group. My hypothesis is that there would be more tickets issued during weekdays than weekends.

```{r manipulate data}
# count number of tickets by weekday
count_by_weekday <-
    tickets %>%
    select(issue_datetime) %>%
    mutate(weekday = wday(issue_datetime, label=TRUE)) %>%
    count(weekday) %>%
    group_by(weekday)
```

## Plot

```{r plot}
# plot number of tickets by weekday
count_by_weekday %>%
    ggplot() +
    aes(x = weekday, y = n, color = weekday, fill = weekday) +
    geom_bar(stat = "identity") + 
    labs(x = "Days of week", y = "Number of tickets") + 
    ggtitle("Number of Tickets by Weekday") +
    theme(legend.position = "none")
```

It turns out my hypothesis was correct. However, I was surprised by how much greater the number of tickets were issued on weekdays versus weekends. Specifically, 3x fewer tickets were issued on Sunday than any other day! 

## Save Plot

```{r save plot}
# save image
ggsave("plot/plot_2019-12-03.png", width = 29, height = 21, units = "cm", dpi = "retina")
```

## Session Info

```{r session info}
#session info
sessionInfo()
```

