---
title: "Hotels"
author: "Francis Lin | #TidyTuesday |"
date: "2020-02-11"
output: rmarkdown::github_document
---

# Introduction

[Antonio, Almeida and Nunes, 2019](https://www.sciencedirect.com/science/article/pii/S2352340918315191#f0010) has provided some hotel data for #TidyTuesday. While looking at this dataset, I noticed that a large amount of datapoints were reservations that were eventually cancelled. I decided to focus my efforts on seeing if I can tease out any trends that I see.

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(readr)
library(lubridate)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```

## Load Data

```{r load data, message=FALSE}
hotels <- read_csv(paste0(data_path, "hotels.csv"))
```

## Manipulate Data

Here, I used the `lubridate` package to derive the scheduled arrival date from the data, and find the difference in days between the cancellation date and the scheduled arrival date. I then wanted to see how the deposit type of the reservation affected the cancellation rates, so I turned that into factors.

```{r manipulate data}
cancels <- hotels %>%
    filter(reservation_status=="Canceled") %>%
    mutate(arrival_date = ymd(paste(arrival_date_year, arrival_date_month, arrival_date_day_of_month, sep="-"))) %>%
    mutate(cancel_lead_time = as.numeric(arrival_date-ymd(reservation_status_date))) %>%
    mutate(deposit = factor(ifelse(deposit_type == "No Deposit", "No Deposit", "Deposit"), levels=c("Deposit", "No Deposit")))
```

## Plot Data

Here, I plotted the number of cancellations per cancellation lead time.

```{r plot data, fig.width=8, fig.height=5}
p <- ggplot(cancels %>% filter(cancel_lead_time < 51)) +
    geom_bar(aes(x=cancel_lead_time, fill=deposit)) + 
    labs(title="Hotel Cancellations", 
         subtitle="Last-minute hotel cancellations occur more frequently, and tend to have no deposit",
         x="Days between Cancellation and Scheduled Arrival Date", 
         y="Number of Cancellations") + 
    theme(legend.title=element_blank())
p
```

From the plot above, it seems the closer you get to the scheduled arrival date, the more cancellations you receive, although it tapers off to the right less drastically than I would have expected. As cancellation lead time increases, the number of cancellation with no deposits increase while the number of cancellation with deposits decrease.

## Save Image

```{r save image}
ggsave("plot/plot_2020-02-11.png", p, width=8, height=5, units="in")
```

## Session Info

```{r session info}
sessionInfo()
```
