---
title: "Rap Artists"
author: "Francis Lin | #TidyTuesday |"
date: "2020-04-14"
output: rmarkdown::github_document
---

# Introduction

[BBC Music](http://www.bbc.com/culture/story/20191007-the-greatest-hip-hop-songs-of-all-time-who-voted) asked multiple critics what their top 5 hip-hop tracks were. Based on the answers, they then awarded points to each song. The analysis below visualizes their responses.

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(ggrepel)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
rankings <- read_csv(paste0(data_path, "rankings.csv"))
```


## Plot Data

```{r plot data, fig.width=8, fig.height=5}
p <- ggplot(rankings, aes(x=year, y=points, color=gender)) +
    scale_color_manual(values = c("female"="red", "male"="royalblue", "mixed"="purple")) +
    geom_point(alpha=0.25) + 
    geom_text_repel(size = 2.5, box.padding = 0.5, aes(label=ifelse(points>as.numeric(quantile(points,0.95)),paste(title, "\n-", artist),''))) + 
    labs(title="Rap Artists and Critic Ratings", x="Year", y="Critic Points", color="Gender")
p
```

The top 5% rap songs with the highest points given by critics between 1980 to 2019 have all been male artists, even though `r format(sum(rankings['gender']!='male')/nrow(rankings)*100, digits=4)`% of the songs are by female or mixed gender artists. "Juicy" by The Notorious B.I.G. was a huge favorite.

## Save Image

```{r save image}
ggsave("plot/plot_2020-04-14.png", p, width=8, height=5, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
