---
title: "GDPR Fines"
author: "Francis Lin | #TidyTuesday |"
date: "2020-04-21"
output: rmarkdown::github_document
---

# Introduction


According to [Wikipedia](https://en.wikipedia.org/wiki/General_Data_Protection_Regulation), The General Data Protection Regulation (EU) 2016/679 (GDPR) is a regulation in EU law on data protection and privacy in the European Union (EU) and the European Economic Area (EEA). In this analysis, I wanted to see which articles result in the highest amounts of fines.

# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
gdpr_text <- read_tsv(paste0(data_path, "gdpr_text.tsv"))
gdpr_violations <- read_tsv(paste0(data_path, "gdpr_violations.tsv"))
```


## Manipulate data

A lot of this work was done by [Julia Silge](https://juliasilge.com/blog/gdpr-violations/).

```{r manipulate data}
gdpr_tidy <- gdpr_violations %>%
  distinct() %>%
  mutate(article = str_extract_all(article_violated, "Art.[:digit:]+|Art. [:digit:]+")) %>%
  mutate(total_articles = map_int(article, length)) %>% 
  mutate(avg_price = price/total_articles) %>%
  unnest(article) %>%
  add_count(article) %>%
  filter(n > 10) %>%
  select(-n) %>%
  select(article, avg_price) %>%
  mutate(article = fct_reorder(article, avg_price))
```


## Plot Data

```{r plot data, fig.width=7, fig.height=5}
p <- ggplot(gdpr_tidy) +
  geom_boxplot(aes(x=article, y=avg_price)) + 
  scale_y_log10() + 
  labs(x="Article", y="Fine (EUR)", title="GDPR Fines")
p
```

Article 32 resulted in the highest average fine amount. You can find the text of the regulation [here](https://www.privacy-regulation.eu/en/article-32-security-of-processing-GDPR.htm).

## Save Image

```{r save image}
ggsave("plot/plot_2020-04-21.png", p, width=7, height=5, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
