---
title: "Uncanny X-men"
author: "Francis Lin | #TidyTuesday |"
date: "2020-06-30"
output: rmarkdown::github_document
---

# Introduction

This week's data comes from the [Claremont Run Project](http://www.claremontrun.com/). It was actually pretty interesting to take a look at this dataset since I actually taught a course on superheroes before! If you're interested, you can find some of the material [here](https://francislin96.github.io/projects/USIE/usie.html). 

Comic series, especially those with superpower themes, are often criticized for appealing to male fantasy and presenting unrealistic body standards. For this week's analysis, I wanted to take a look to see how often issues from the X-men comic series pass the Bechdel test. My hope is that passing the Bechdel test becomes more frequent as the issue number increases.


# R Program

## Set up

```{r setup, message=FALSE}
# source utilities
source("../utilities.R")

# packages
library(tidyverse)
library(gridExtra)
library(grid)

# set data path
data_path <- set_data_path()

# create plot folder if doesn't exist
if (!dir.exists("./plot")) {dir.create("./plot")}
```


## Load Data

```{r load data, message=FALSE}
comic_bechdel <- read_csv(paste0(data_path, "comic_bechdel.csv"))
xmen_bechdel <- read_csv(paste0(data_path, "xmen_bechdel.csv"))
```


## Manipulate data

```{r manipulate data, message=FALSE}
factor_to_string <- function(fctr){
    split <- strsplit(gsub("\\[|\\(|\\]|\\)", "", fctr), ",")[[1]]
    return(paste0(as.character(as.numeric(split[1])+1), "-", split[2]))
}

bechdel_comic <- comic_bechdel %>%
    
    # select relevant columns and dropna
    select(issue, pass_bechdel) %>%
    drop_na(issue, pass_bechdel) %>%
    
    # recode
    mutate(pass_bechdel = recode(pass_bechdel, yes = "Pass", no = "Fail")) %>%
    
    # factorize issue column
    mutate(issue_factor = cut(x = issue, breaks = seq(0, 300, 10))) %>%
    arrange(issue_factor) %>%
    #mutate(issue_factor = sapply(issue_factor, factor_to_string)) %>%
    
    # for each issue factor, calculate how many pass/fail bechdel test
    group_by(issue_factor, pass_bechdel) %>%
    summarize(count=n()) %>%
    ungroup() %>%
    group_by(issue_factor) %>%
    mutate(percent_pass = count/sum(count))


bechdel_xmen <- xmen_bechdel %>%
    
    # select relevant columns and dropna
    select(issue, pass_bechdel) %>%
    drop_na(issue, pass_bechdel) %>%
    
    # recode
    mutate(pass_bechdel = recode(pass_bechdel, yes = "Pass", no = "Fail")) %>%
    
    # factorize issue column
    mutate(issue_factor = cut(x = issue, breaks = seq(0, 550, 20))) %>%
    arrange(issue_factor) %>%
    #mutate(issue_factor = sapply(issue_factor, factor_to_string)) %>%
    
    # for each issue factor, calculate how many pass/fail bechdel test
    group_by(issue_factor, pass_bechdel) %>%
    summarize(count=n()) %>%
    ungroup() %>%
    group_by(issue_factor) %>%
    mutate(percent_pass = count/sum(count))
```


## Plot Data

```{r plot data, fig.width=12, fig.height=8}
p1 <- ggplot(bechdel_comic, aes(fill=pass_bechdel, y=count, x=issue_factor)) + 
    
    # bar chart
    geom_bar(position="fill", stat="identity") + 
    
    # label
    geom_text(aes(label = ifelse(pass_bechdel=="Pass", paste0(round(100*percent_pass),"%"),""),y=percent_pass-0.03),size = 3) +
    
    # relabel x axis
    scale_x_discrete(labels=sapply(levels(bechdel_comic$issue_factor), factor_to_string)) +
    
    # aesthetics
    theme_minimal() +
    theme(
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(vjust = -1, size=18)
        ) +
    labs(x="Issue", y="Percentage", fill="", title="Non-X-Men Comic Series")

p2 <- ggplot(bechdel_xmen, aes(fill=pass_bechdel, y=count, x=issue_factor)) + 
    
    # bar chart
    geom_bar(position="fill", stat="identity") + 
    
    # label
    geom_text(aes(label = ifelse(pass_bechdel=="Pass", paste0(round(100*percent_pass),"%"),""),y=percent_pass-0.03),size = 3) +
    
    # relabel x axis
    scale_x_discrete(labels=sapply(levels(bechdel_xmen$issue_factor), factor_to_string)) +
    
    # aesthetics
    theme_minimal() +
    theme(
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.grid.major = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(vjust = -1, size=18)
        ) +
    labs(x="Issue", y="Percentage", fill="", title="Uncanny X-Men Comic Series")

p <- grid.arrange(p2, p1, nrow=2, top=textGrob("Comic Series Bechdel Test Results", gp=gpar(fontsize=24)))
```


## Save Image

```{r save image}
ggsave("plot/plot_2020-06-30.png", p, width=12, height=8, units="in")
```


## Session Info

```{r session info}
sessionInfo()
```
